---
title: "Data Tidying Lecture Code and Questions"
output: html_document
editor_options: 
  chunk_output_type: console
---

Adapted from UO Psych R Bootcamp: <https://uopsych-r-bootcamp-2020.netlify.app/post/08-tidyr/>

## Instructions

- Open the visual markdown editor.
- Run the code chunk below. You may need to `install.packages("nycflights13")` first.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)

flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

```

---

**left_join example 1**

```{r}
flights2 %>%
  left_join(airlines, by = "carrier")
```

**left_join equivalent using base R subsetting**

```{r}
tibble(flights2,
       name = airlines$name[match(flights2$carrier, airlines$carrier)])
```

---

**inner_join example 1**

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     3, "x3"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     4, "y3"
)

x %>% 
  inner_join(y, by = "key")
```

---

**left_join example 2**

```{r}
    x <- tribble(
      ~key, ~val_x,
         1, "x1",
         2, "x2",
         2, "x3",
         1, "x4"
    )
    y <- tribble(
      ~key, ~val_y,
         1, "y1",
         2, "y2"
    )
    left_join(x, y, by = "key")
```

---

**left_join example 3**

```{r}
flights2 %>% 
      left_join(weather)
```

---

**left_join example 4**

```{r}
flights2 %>% 
  left_join(planes, by = "tailnum")
```

---

*left_join example 5*

Map destination airport (dest) to faa:

```{r}
flights2 %>% 
  left_join(airports, c("dest" = "faa"))
```



Map origin airport to faa:

```{r}
flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```

---

*semi_join example 1*

```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)

top_dest
```


```{r}
flights %>% 
  filter(dest %in% top_dest$dest)
```


```{r}
flights %>% 
  semi_join(top_dest)
```


```{r}
flights %>% 
  semi_join(top_dest)
```

---

**anti_join example 1**

```{r}
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)
```

---

**Set operations example**

```{r}
df1 <- tribble(
  ~x, ~y,
   1,  1,
   2,  1
)
df2 <- tribble(
  ~x, ~y,
   1,  1,
   1,  2
)

df1

df2
```

```{r}
intersect(df1, df2)

union(df1, df2)

setdiff(df1, df2)

setdiff(df2, df1)
```

---

**pivot_longer()**

```{r}
table4a %>% 
  pivot_longer(cols = c(`1999`, `2000`),
               names_to = "year", 
               values_to = "cases")
```

---

**Your Turn 1**

```{r}
# Q1.

ps_data <- rio::import("data/pragmatic_scales_data.csv") %>% 
  as_tibble()
```


```{r}
#Q2.

```


```{r}
#Q3.

```

---

**separate and unite example 1**

```{r}
table3
```


```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"))
```


```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
```


```{r}
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```


```{r}
table5 %>% 
  unite(new, century, year)
```


```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```

---

**Missing values example 1**

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)

stocks
```


```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return)
```


```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(
    cols = c(`2015`, `2016`), 
    names_to = "year", 
    values_to = "return", 
    values_drop_na = TRUE
  )
```

---

**Missing values example 2**

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)

treatment
```


```{r}
stocks %>% 
  complete(year, qtr)
```
